---
title: "dplyr"
author: "virgile-baudrot"
date: "Wednesday, October 29, 2014"
output: html_document
---

**dplyr** is made of 5+1 functions:
- filter: keep rows matching criteria
- select: pick columns by name
- arrange: reorder rows
- mutate: add new variables
- summarise: reduce variables to values
and the final function:
- group_by: used to group data according to criteria

This document is used to present the package dyplr. 

```{r}
#load packages
library('dplyr')
library('hflights')

# explore data
data(hflights)
head(hflights)
```

'tbl_df() is a function creating a local data frame

```{r}
flights<-tbl_df(hflights)
flights
```

print() may be used to print a head with a greater number of row of the local data frame object

```{r}
print(flights,n=20)
```


## filter : choose specific lines

Native function to view flight of the first day of january

```{r}
flights[flights$Month==1 & flights$DayofMonth==1,]
````

with dplyr, the functoion filter() is more intuitive (easy to understand)

```{r}
filter(flights, Month==1, DayofMonth==1)
#
filter(flights, UniqueCarrier=="AA" | UniqueCarrier=="UA")
filter(flights, UniqueCarrier %in% c("AA","UA"))
```


## select : pick columns

The classical R way to do it:

```{r}
flights[,c("DepTime","ArrTime","FlightNum")]
```

and the dplyr way:

```{r}
select(flights, DepTime,ArrTime,FlightNum)
```

## Chaining

Chaining is probably the more inovative function in the package.
See the nesting method:

```{r}
# select UniqueCarrier and DepDelay columns and fliter for delays over 60 minutes:

filter(select(flights,UniqueCarrier, DepDelay),DepDelay>60)
```

the chaining way increases readability:

```{r}
flights %>% # select the object
  select(UniqueCarrier, DepDelay) %>% #apply the first function
  filter(DepDelay >60) # and the second function
```

Create two vectore to calculate euclidian distance detween them:

```{r}
x1<-1:5
x2<-2:6
# the chaining method:

(x1-x2)^2 %>% sum() %>% sqrt()
```


## arrange : reorder row

select UniqueCarrier and DepDelay columns and sort by DepDelay

```{r}
flights %>%
  select(UniqueCarrier, DepDelay) %>%
  arrange(desc(DepDelay))
```

## mutate: add new variables

creation of a new variable speed (in mph)

```{r}
flights %>%
  select(Distance, AirTime) %>%
  mutate(Speed = Distance/AirTime*60)
```

## summarise: reduce variables to values

create a table grouped by Dest, and then summarise each groupe by taking the mean of ArrDalay

```{r}
flights %>%
  group_by(Dest) %>%
  summarise(avg_delay =mean(ArrDelay,na.rm=TRUE))
```


## summarise_each: apply the same function to multiple columns at once

for each carrier, calculate the percentage of flights cancelled or diverted

```{r}
flights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(mean),Cancelled, Diverted)
```

for each carrier, calculate the minimum and maximum arrival and departure delays

```{r}
flights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(min(.,an.rm=TRUE),max(.,na.rm=TRUE)),matches("Delay"))
```

## Example of %>% the pipeplines operator:

Another set of data

```{r}

```


```{r}
hourly_delay <- flights %>%
  filter(!is.na(dep_delay)) %>%
  group_by(date,hour) %>%
  summarise(
    delay=mean(dep_delay),
    n=n()
    ) %>%
  filter(n>10)

```



