---
title: "ggplot2"
author: "virgile-baudrot"
date: "Thursday, October 30, 2014"
output:
  html_document:
    highlight: tango
    theme: flatly
---

Beginning
----
## A plot laye by layer

size, colour and shape is *aesthetics*. ggplot are like ogre (Shrek), it's made of layers! And so like oignon. Layers are:
- data and aesthetic,
- statistical transformed,
- geometric object,
- position adjustement


First, load the library ggplot2, and see the data diamonds (include in ggplot2). A dataset containing the prices and other attributes of 54000 diamonds.

```{r}
library(ggplot2)
head(diamonds)
````

We can used a random sample of the dataset diamonds:
```{r}
set.seed(1410) # make the sample repoducible
dsmall<-diamonds[sample(nrow(diamonds),100),]
head(dsmall)
```

First look at the data:

```{r}
#a simple plot of all the diamonds
qplot(carat,price,data=diamonds)
# another plot of the selection
qplot(carat,price,data=dsmall)
```

```{r}
p <- ggplot(diamonds, aes(carat,price,colour=cut))
```

In the object *p*, there is no layer. Adding the layer is by: 
 
```{r}
p<-p+layer(geom="point")
p
```

A stat or geom layer is made by the same expression:
- geom_XXX(...)
- stat_XXX(...)

example of $qplot()* with a layer:
```{r}
# data
head(msleep)
# without layer:
qplot(sleep_rem/sleep_total, awake,data=msleep)
#and with a layer:
qplot(sleep_rem/sleep_total, awake,data=msleep) + geom_smooth()
```

layer with *ggplot()*

```{r}
# the same with layer:
p<-ggplot(msleep,aes(sleep_rem/sleep_total,awake)) + geom_point() + geom_smooth()
p
# the summary function:
summary(p)
```


Interest of layers?
```{r}
library(scales) # to make transparency
# this is how you what the curve fit look like:
bestfit <- geom_smooth(method="lm",se=F,colour=alpha("steelblue",0.5),size=2)
#and the data you use:
qplot(sleep_rem,sleep_total,data=msleep)+bestfit
# another dataset:
qplot(bodywt,brainwt,data=msleep,log="xy") + bestfit
```


## Aesthetic mapping



## Transparency

Transparency need the packages *scales* in order to used the function *alpha()*

```{r}
df<-data.frame(x_norm=rnorm(2000),y_norm=rnorm(2000))
library(ggplot2)
norm <- ggplot(df, aes(x=x_norm,y=y_norm))
norm+geom_point()
norm+geom_point(shape=1)
norm+geom_point(shape=".")
### Transparency
library(scales)
norm+geom_point(colour=alpha("black",1/3))
norm+geom_point(colour=alpha("black",1/10))
```

## Annotating a plot

we'll use data economics already load in the package *ggplot2*

```{r}
library(ggplot2)
# data economics are already in the package
head(economics)
unemp<-qplot(date,unemploy,data=economics, geom="line")
unemp
# adding legent on qplot
unemp<-qplot(date,unemploy,data=economics, geom="line",
             xlab="",ylab="No. unemployed(1000s)")
unemp
```

### geom_vline()

```{r}
# data presidential  are already in the package
presid <- presidential[-(1:3),]
presid
# range: return vector with minimum and maximum
yrng <- range(economics$unemploy)
yrng
xrng<-range(economics$date)
xrng
unemp+geom_vline(aes(xintercept=as.numeric(start)),data=presid)

```


### geom_vline()

```{r}
unemp + 
  geom_rect(aes(NULL, NULL, 
                xmin = start, xmax = end, 
                fill = party), 
            alpha = 0.2, 
            ymin = yrng[1], ymax = yrng[2], 
            data = presid) + 
  scale_fill_manual(values = c("Republican"="red", 
                               "Democratic"="blue"))
```

### last_plot()

```{r}
last_plot() + 
  geom_text(aes(x = start, y = yrng[1], label = name), 
            data = presid, size = 3, hjust = 0, vjust = 0)
```

