---
title: "GoogleVis"
author: "virgile-baudrot"
date: "Friday, November 07, 2014"
output: html_document
---

Here, I'll present the package googleVis which provides an interface to google charts API. From Geo Charts to Scatter Charts and so on.
- R create an HTML,
- The HTML page call google charts,
- HTML interactive graphics

```{r}
library(googleVis)
```


What is it use for ?

```{r}
#demo(googleVis) 
# then click enter to see a graph open in your browser...
```

To plot in the r markdown, we have to add *results="asis"* in the chunk options

```{r,results='asis'}
data(Fruits)
Fruits
M<-gvisMotionChart(Fruits,idvar="Fruit",timevar="Year")
str(M)
M$type
M$chartid
print(M,tag='header')
names(M$html$chart)
```

To have the complete chart displayed:
```{r,results='asis'}
print(M,tag="chart")
```

and a beautiful plot open in the browser

```{r,results='asis'}
plot(M)
```

we can create an HTML file to save in the directory

```{r}
print(M, file="myGoogleVisChart.html")
```


and we can open it with:

```{r}
#plot.gvis(file="myGoogleVisChart.html")
```


```{r}
#suppressPackageStartupMessages(library(googleVis))
M <- gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=600, height=400))
print(M,"chart")
```

## Geographic charts

```{r,results='asis'}
G <- gvisGeoChart(Exports, locationvar="Country",
                  colorvar="Profit",options=list(width=600, height=400))
print(G,"chart")
plot(G)
```



```{r, results='asis'}
G2 <- gvisGeoChart(Exports, locationvar="Country",
                  colorvar="Profit",options=list(width=600, height=400,region="150"))
print(G2,"chart")
plot(G2)
```


## Other data

```{r}
df <- data.frame(label=c("US", "GB", "BR"), val1=c(1,3,4), val2=c(23,12,32))
Line <- gvisLineChart(df, xvar="label", yvar=c("val1","val2"),
        options=list(title="Hello World", legend="bottom",
                titleTextStyle="{color:'red', fontSize:18}",                         
                vAxis="{gridlines:{color:'red', count:3}}",
                hAxis="{title:'My Label', titleTextStyle:{color:'blue'}}",
                series="[{color:'green', targetAxisIndex: 0}, 
                         {color: 'blue',targetAxisIndex:1}]",
                vAxes="[{title:'Value 1 (%)', format:'##,######%'}, 
                                  {title:'Value 2 (\U00A3)'}]",                          
                curveType="function", width=500, height=300                         
                ))
```


```{r,results='asis'}
G <- gvisGeoChart(Exports, "Country", "Profit",options=list(width=200, height=100))
T1 <- gvisTable(Exports,options=list(width=200, height=270))
M <- gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=400, height=370))
GT <- gvisMerge(G,T1, horizontal=FALSE)
GTM <- gvisMerge(GT, M, horizontal=TRUE,tableOptions="bgcolor=\"#CCCCCC\" cellspacing=10")
plot(GTM)
```


```{r,results='asis'}
M <- gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=600, height=400))
print(M)
plot(M)
#print(M, 'chart', file='myfilename.html')
```


