---
title: "Loops"
author: "virgile-baudrot"
date: "Monday, November 10, 2014"
output: html_document
---

Here is a lecture for loops.

## My first loop **for**


#### What is a loop ?


```{r}
for(i in 1:10){
  print(i)
}
```

#### A loop is very long in R

Let consider these data:

```{r}
n<-1e6 # a big number
g<-rnorm(n)
h<-rep(NA,n)
```


#### Timing code

A loop adding 1 to every element of g:

```{r}
# start the clock
ptm<-proc.time()
for(i in 1:n){
  h[i]<-g[i]+1
}
# stop the clock
loop.time<-proc.time()-ptm
```

The *user time* is the CPU time charged for the execution of user instructions of the calling process. The *system time* is the CPU time charged for execution by the system on behalf of the calling process.

- user (utilisateur): the execution of the code,
- system (système): time relate to CPU
- elapsed (écoulé) sum of user and system if all lines are run together

To avoid start and stop code, we can use `system.time` function:

```{r}
# wrapping the code in a function named loop.code
loop.code<-function(g){
  h<-rep(NA,length(g))
  for(i in 1:n){
    h[i]<-g[i]+1
  }
}
# time to compute:
system.time(loop.code(g))
```

Without a loop (in a vector approach) the process is faster:

```{r}
vector.code<-function(g){
    g+1
}
system.time(vector.code(g))
```

This is just super fast!


## apply, sapply, lapply, tapply,...


apply: lets take a big matrix with values of a normal law with mean 3 and variance 9 (sd=3)

```{r}
m<-matrix(rnorm(1e7,3,3),nrow=1e2) # so we have 1e4 column: try ncol(m)
print(object.size(m),units="GB")
```

And we want the mean of each column:

```{r}
#with a loop
mean.col<-function(m){
  mean.col.vec<-rep(NA,ncol(m))
  for(i in 1:ncol(m)){
    mean.col.vec[i]<-mean(m[,i])
  }
  return(mean.col.vec)
}
system.time(mean.col(m))
# with apply
system.time(apply(m,2,mean))
```

and with a bigger matrix:

```{r}
m.big<-matrix(rnorm(1e8,3,3),nrow=1e3) # so we have 1e4 column: try ncol(m)
print(object.size(m.big),units="GB")
```

And we want the mean of each column:

```{r}
#with a loop
system.time(mean.col(m.big))
# with apply
system.time(apply(m.big,2,mean))

```

The function `apply()` is not always faster...

There is a function a lot faster:

```{r}
system.time(colMeans(m.big))
rm(list=ls()) # remove other object
```



## My first if

If you have more than `if`.


```

## Cleaning the 

```{r}
rm(list=ls()) # enlever ce qui est dans l'espace de travail
```