---
title: "R-Meetup"
author: "Virgile Baudrot"
date: "Thursday, November 13, 2014"
output: ioslides_presentation
runtime: shiny
---

## Ready for a great presentation?

Présentation avec slidify.

```{r,echo=FALSE}
inputPanel(
  selectInput("l_level", label = "Select a word",
              choices = c("R c'est cool", "R c'est utile","R c'est naze"), selected ="R c'est utile")
)
```

And the result is:

```{r,echo=FALSE}
renderText({
  if(input$l_level=="R c'est utile"){
    "Parfois"
  }
  else if(input$l_level=="R c'est naze"){
    "Oh non eh !!!"
  }
  else "Yes gros"
})
```


## Shiny Presentation

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately. 

To learn more, see [Interative Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## What I'll talk about:

- What is a loop?
- Why R studio is so good?
- Why markdown is very usefull?
- Example: How to create a package in 5 minutes?
- How I did this shiny presentation


## Slide with Interactive Plot

A data set 

```{r,echo=FALSE}
inputPanel(
  sliderInput("nb_people", label = "Nombre de participant",
              min = 1, max = 30, value = 20, step = 1),
  sliderInput("nb_lovR", label = "Nombre de LoveR",
              min = 1, max = 30, value = 10, step = 1),  
  sliderInput("nb_needR", label = "Nombre de needR",
              min = 1, max = 30, value = 10, step = 1)   
)
```

```{r,echo=FALSE}
renderText({
   print(c("Nombre des autres : ",input$nb_people - (input$nb_lovR+input$nb_needR)))
})
```

```{r,echo=FALSE,warning=FALSE}
library(ggplot2)
```

```{r,echo=FALSE}
renderPlot({
  df=data.frame(x=input$nb_needR,y=input$nb_lovR)
  ggplot(df,aes(x=x,y=y)) +ylim(c(0,30))+xlim(c(0,30))+
    xlab("Nombre de lovR")+ ylab("Nombre de needR") +
    geom_point(colour="red",size=3)
})
```


## 2nd slide with interactive plot

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```


## My first loop:

```{r}
for(i in 1:10){
  print(i)
}
```

## A loop is very long...

Let consider these data:

```{r}
n<-1e6 # a big number
g<-rnorm(n) # 1 million randomlly data following normal(0,1)
h<-rep(NA,n) #a vector of length 1 million
```

How long is a loop:

```{r}
# start the clock
ptm<-proc.time() # register the time
for(i in 1:n){
  h[i]<-g[i]+1 # my function
}
# stop the clock
proc.time()-ptm
```


## system.time()

```{r}
# wrapping the code in a function named loop.code
loop.code<-function(g){
  h<-rep(NA,length(g))
  for(i in 1:n){
    h[i]<-g[i]+1
  }
}
# time to compute:
(system.time(loop.code(g)))
```


## R STUDIO

Let see it in live !

## R Markdown

Let see it in live !

## A package in 5 minutes!


